Comparing tool variability and numerical variability in fMRI analyses

### 2021-10-12 paper/paper.tex
Comments in results

@@

For our analyses, we used the Fuzzy libmath (FL) library introduced in~\cite{salari2021accurate}, an MCA-instrumented 
version of the mathematical library in GNU operating systems (libmath).
Using FL, we can simulate numerical variability at the level of machine error, the floating-point rounding error
caused by floating-point representation limits defined by the hardware.

@@

To avoid some pitfalls in deterministic arithmetics like producing results out of the function definition,
FL instruments the functions by wrapping them so that only the original functions' output
is perturbed instead of input or their implementation.
Listing~\ref{algo:wrapper} shows an example of this wrapping for the log function in both single and double precisions.
In this script, the original functions are called by \texttt{dlsym},
a function that returns the memory address of a symbol using the handle of \texttt{RTLD\_NEXT}.
This allows one to provide a wrapper around a function in another shared library.
Also, the MCA instrumentation is activated by adding a floating-point zero to the output of the original function
so that the result of this summation is perturbed and returned.

@@

FL allows measuring the effect of numerical variability by running a program multiple times
with the feature of controlling the magnitude of the perturbations as perceived by the application.
FL enables us to assess the software packages that are dynamically linked to the mathematical library.
So, it is important to trace the tool dependencies of the pipelines to understand the different libraries
involved during the pipeline executions.

We replicated the fMRI analysis described as study `ds000001' in~\cite{bowring2019exploring}, relying on
the data publicly available in OpenNeuro at \url{https://openneuro.org/datasets/ds000001}
and using the three main software packages for fMRI data processing, namely
FMRIB Software Library (FSL)~\cite{jenkinson2012fsl}, Analysis of Functional NeuroImages (AFNI)~\cite{cox1996afni},
and Statistical Parametric Mapping(SPM)~\cite{penny2011statistical}.
We selected this analysis because the methods and datasets were explicitly
written to be accessible and reproducible. This enabled us to reanalyze fMRI experiments using the FL framework and
compare WT and BT variabilities.

In the selected study, 16 healthy adult subjects participated in the balloon analog risk task to measure
risk-taking behavior over three scanning sessions.
This study included preprocessing, first-level, and second-level analyses that were implemented with all three software.
Table~\ref{table:pipeline-steps} shows the steps that were taken in the procedure of analyses for each tool in the original study.
In the original study, a number of preprocessing steps widely accepted within the community, such as motion correction,
segmentation, brain extraction, and registration, were applied in all analyses to ensure that results from each software
package could be compared objectively.
A full description of the pipelines implemented within three packages is presented in~\cite{bowring2019exploring,schonberg2012decreasing}.

      \end{tabular}
    \caption{Software processing steps. Table is adapted from~\cite{bowring2019exploring}.}
    \label{table:pipeline-steps}
\end{table}

\subsection{Data processing}

To capture variability in BT, the fMRI analysis pipeline was processed using three of the most popular software
packages in neuroimaging, including FSL (version 5.0.10), AFNI (version 18.1.09), and SPM (version SPM12, r7771)
executed with GNU/Octave (version 5.2). 
FSL and AFNI scripts were executed in Python 2.7. We also used a fixed number of 7 threads in AFNI executions
to reproduce similar results to the original study in a feasible execution time.
All the analyses were conducted on the same operating system, CentOS 7.3.
We ensured that the software versions used in all experiments were similar to the original study,
and encapsulated them in Docker container images with available Dockerfiles at \url{https://github.com/ali4006/fuzzy-neurotools/tree/main/dockerfile}.

In addition, the same subjects and fMRI analysis were processed with the same configurations in the FL environment
to produce results in WT variability.
We applied instrumentations at the virtual precision (t) of 53 bits for double-precision floating-point values
and 24 bits for single-precision values. Three Fuzzy samples were generated, to match the number of tool samples.

We evaluated WT and BT variabilities in thresholded and unthresholded group-level activation maps.
We computed voxel-wise standard deviations of T-statistic values for each pair of tools
and then compared the distribution of standard deviations in both conditions.
We computed the WT standard deviation as the square root of the summation of variances between samples in each tool.
Moreover, we determined region-by-region Dice coefficients for the thresholded maps for each pair of tools.
We characterized 360 regions that correspond to the cortical parcellation atlas
in Human Connectome Project Multi-Modal Parcellation version 1.0 (HCP-MMP1.0)~\cite{glasser2016multi}.
Dice values in WT are computed as the average pair-wise Dices among three Fuzzy samples for each pair of tools.
Also, we compared the correlation of Dice scores normalized by the region sizes in both conditions.

>>>>

We replicated the fMRI analysis described as study `ds000001'
in~\cite{bowring2019exploring} \tristan{check if this dataset was acquired
and analysed previously, and if so cite the original paper}, relying on the
data publicly available in OpenNeuro at
\url{https://openneuro.org/datasets/ds000001} and using the three main
software packages for fMRI data processing, namely FMRIB Software Library
(FSL)~\cite{jenkinson2012fsl}, Analysis of Functional NeuroImages
(AFNI)~\cite{cox1996afni}, and Statistical Parametric
Mapping(SPM)~\cite{penny2011statistical}. We selected this dataset because
comparable analysis pipelines implemented in FSL, AFNI and SPM 
were already published and extensively described in~\cite{bowring2019exploring}.
Furthermore, the work in~\cite{bowring2019exploring} already evaluated the effect of tool variability for
this dataset, which we intended to complement with the present quantification of numerical variability.

In the selected study, 16 healthy adult subjects participated in the
balloon analog risk task \tristan{cite the task specifically} to measure
risk-taking behavior over three scanning sessions \tristan{cite the
original study (not Alex')}. We reused the preprocessing, first-level, and
second-level analyses implemented by~\cite{bowring2019exploring} consistently across all three software packages using 
widely accepted analytical steps. Table~\ref{table:pipeline-steps} summarizes the analytical steps in each pipeline.
More details are available in~\cite{bowring2019exploring,schonberg2012decreasing} \tristan{cite only one ref, the one that gives the right details}.

      \end{tabular}
    \caption{Software processing steps (adapted from~\cite{bowring2019exploring}).}
    \label{table:pipeline-steps}
\end{table}

@@

To introduce controlled amounts of numerical error in the analyses, we used the Fuzzy Libmath (FL) library~\cite{salari2021accurate}, an MCA-instrumented 
version of the GNU mathematical library (libmath).
Using FL, we can simulate numerical variability at the level of machine error \tristan{you really shouldn't put so much emphasis on machine error so early in the description. First, explain how FL works, then, say that you used it to simulate machine error but not only.}, the floating-point rounding error
caused by floating-point representation limits defined by the hardware \tristan{this is a quite bad definition of machine error}.

@@

FL only introduces perturbations in the output values of mathematical
functions but not in their implementation. This is done by wrapping the original functions 
and applying function \texttt{inexact} to their returned values.
Listing~\ref{algo:wrapper} shows an example of this wrapping for the log function in both single and double precisions.
In this script, the original functions are called by \texttt{dlsym},
a function that returns the memory address of a symbol, in our case \texttt{RTLD\_NEXT}, the address of the next occurrence of the function.
These wrappers are compiled with Verificarlo for each function. The MCA
instrumentation is activated by adding a floating-point zero to the output
of the original function so that this summation is instrumented.

@@

FL allows measuring the effect of numerical variability by running a program multiple times, 
resulting in multiple independent realizations of the same analysis. Each of these samples 
are equally plausible estimates of the true numerical result.

FL enables us to assess the software packages that are dynamically linked to the mathematical library.
So, it is important to trace the tool dependencies of the pipelines to understand the different libraries
involved during the pipeline executions \tristan{this is not methods. Here you should report how you validated the instrumentation 
of each pipeline: (1) dependency analysis, (2) runtime checks, etc}.
\subsection{Data processing}

To measure between-tool (BT) variability, we executed the pipelines described in~\cite{bowring2019exploring}
with FSL version 5.0.10, AFNI version 18.1.09, and SPM12 version r7771
executed with GNU/Octave version 5.2. \tristan{are these versions the same as in Alex' paper?} 
FSL and AFNI scripts were executed in Python 2.7 \tristan{why?}. We also used a fixed number of 7 threads in AFNI executions,
to reproduce similar results in~\cite{bowring2019exploring} in a feasible execution time.
All the analyses were conducted on the same operating system, CentOS 7.3.
We ensured that the software versions \tristan{you mentioned versions before, the writing is a bit circular here} used in all experiments were similar to the original study,
and encapsulated them in Docker container images with available Dockerfiles at \url{https://github.com/ali4006/fuzzy-neurotools/tree/main/dockerfile}.

The same subjects and analyses were conducted in the same configurations
with the FL environment, to measure within-tool (WT) variability. We
applied instrumentations at the virtual precision t=53 bits for
double-precision floating-point values and t=24 bits for single-precision
values. Three fuzzy samples were generated for each subject and tool, to
match the number of tool samples. \tristan{mention that for FSL you also used more virtual precisions}

We evaluated WT and BT variabilities in thresholded and unthresholded group-level and subject-level t-statistics maps.
We computed voxel-wise standard deviations of T-statistic values for each pair of tools
and then compared the distribution of \tristan{there's no distribution analysis in your results} standard deviations in both conditions.
We computed the WT standard deviation as the square root of the summation of variances between samples in each tool. \tristan{write it as a formula, it's important enough and will be clearer}

Moreover, we determined region-by-region Dice coefficients for the thresholded maps for each pair of tools.
We characterized 360 regions that correspond to the cortical parcellation atlas
in Human Connectome Project Multi-Modal Parcellation version 1.0 (HCP-MMP1.0)~\cite{glasser2016multi}.
Dice values in WT are computed as the average pair-wise Dices among three Fuzzy samples for each pair of tools.
Also, we compared the correlation of Dice scores normalized by the region sizes in both conditions \tristan{move to results}.
